---
title: "Predictive Model for Average CPU Usage"
author: "Brian Kichler"
date: "May 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building the Model

Logistic regression is an appropriate method to predict the probability of an event by defining a dichotomous/dummy variable. In this case, we want to see if we can develop a model that predicts whether a VM will be underutilized and a candidate for reduction in number of CPUs.

We can start by defining a new dummy variable, which will be on AvgCpuUsage in wide_df. We'll create the variable "BelowQ1AvgCpuUsage", which is defined as true (1) when an observation falls below the first quartile value for Average CPU Usage:

```{r}
wide_df_model <- wide_df %>%
  mutate(BelowQ1AvgCpuUsage = as.numeric(AvgCpuUsage < 168236689))
```

There will obviously be bias due to the definition of this variable. Here's what it looks like:

```{r}
table(wide_df_model$BelowQ1AvgCpuUsage)
```

To mitigate this, this code creates a data frame called "trainingData," which equally samples ones and zeroes with a set seed, then folds the leftover data into another data frame called "testData."


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
